The Power of Two Random Choices\\ \large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Träff, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}

### 2014-01-26 article/article.tex
history of 2-choice paradigm

The paper is organized as follows: We will first introduce different variations of the \emph{balls-into-bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}. Sections \ref{sec:single-choice} to \ref{sec:multiple-choice} will deal with the most basic variation of the balls-into-bins problem, namely the sequential, finite and static variant (see Section \ref{sec:classification} for details). For comparison and as a starting point, Section \ref{sec:single-choice} provides the results for the single-choice balls-into-bins problem. \todo{TODO: multiple-choice, analysis, mention that witness tree method very helpful}

@@

\subsubsection{History}
\label{sec:historyOfGreedy}
This was first proven by Azar et al. \cite{ABKU99} for the case $m = n$. However, the case $d = 2$ was implicitly proven by Karp et al. \cite{KLM92}. See also \cite{MRS01} for a simplified proof and an overview of proof techniques for the given problem.  Unfortunately, the analysis breaks down for the heavily loaded case, where $m = \omega\left(n\right)$. Berenbrink at. all \cite{BFZR08} were the first to proof that the same bound holds for the heavily loaded case\footnote{However, they say that an event $\mathcal E$ occurs with high probability if $\Pr\left[\mathcal E \right]  \geq 1- n^{-\alpha}$, for an arbitrarily chosen constant $\alpha \geq 1$.}. A simplified proof may be found in \cite{TW13}.

>>>>

The paper is organized as follows: We will first introduce different variations of the \emph{balls-into-bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}. Sections \ref{sec:single-choice} to \ref{sec:multiple-choice} will deal with the most basic variation of the balls-into-bins problem, namely the sequential, finite and static variant (see Section \ref{sec:classification} for details). For comparison and as a starting point, Section \ref{sec:single-choice} provides the results for the single-choice balls-into-bins problem. \todo{TODO: multiple-choice, analysis, mention that witness tree method very helpful}  See also \cite{MRS01} an overview of proof techniques for the given problem.

@@

\subsubsection{History}
\label{sec:historyOfGreedy}
The history of proving this remarkable property of the two-choice paradigm is in itself quite interesting. It was first proven by Azar et al. \cite{ABKU99} for the case $m = n$. In the same paper, they claimed that the bound holds for arbitrary $m$. Unfortunately, their analysis breaks down for the heavily loaded case, where $m = \omega\left(n\right)$. Berenbrink at. all \cite{BFZR08} were the first to proof that indeed the same bound holds for the heavily loaded case\footnote{However, they use a slightly different definition for w.h.p.: An event $\mathcal E$ occurs with high probability if $\Pr\left[\mathcal E \right]  \geq 1- n^{-\alpha}$, for an arbitrarily chosen constant $\alpha \geq 1$.}. A simplified proof may be found in \cite{TW13}. Curiously, the case $d = 2$ was implicitly proven by Karp et al. in the context of simulating PRAMs on distributed memory machines\cite{KLM92}. As we will see in the next section, V\"ocking was able to slightly improve the bound as well as to show that the always-go-left strategy that he introduces is optimal up to additive constants \cite{VOC03}.
